Intersection of non-matching grids of different dimensions

### 2016-09-26 paper_intersections.tex

\subsection{Intersection Line-Triangle (1D-2D)}
\notePE{In this paragraph we investigate real lines. Trimming comes later when tolerance is denoted, see next PE note.}
Let us consider a line $p$ with parametric equation $\vc X = \vc A + t\vc u$, $t\in\R$ and a triangle given by vertices $(\vc V_0, \vc V_1, \vc V_2)$ 
with oriented sides $s_i=(V_j, V_k)$, $j=(i+1)\text{ mod }3$, $k=(i+2)\text{ mod }3$. The permuted inner products $\pi_p \odot \pi_{s_i},\, i=0,1,2$ have the same sign if and only if there
is an intersection point with the       line $p$ inside the triangle. In such a case, the inner products provides scaled barycentric coordinates of the intersection on the triangle.
In particular for the barycentric coordinate $w_i$ related to the vertex $V_i$ we have
\[
   w_i = \frac{\pi_p \odot \pi_{s_i}}{ \sum_{j=0}^{2} \pi_p \odot \pi_{s_i} }
\]
Indeed, using the barycentric coordinates the intersection point can be expressed as $\vc X = \vc V_0 + w_1 \vc s_2 - w_2 \vc s_1$.
The line $p$ have \plucker coordinates $(\vc u, \vc u \times \vc X)$ (\plucker coordinates are invariant to change of initial point). Combining these two expressions we get
\[
   \pi_p \odot \pi_{s_1} = \vc u \cdot (\vc s_1 \times V_2) + \vc s_1 \cdot ( \vc u \times [\vc V_0 + w_1\vc s_2 - w_2 \vc s_1])
   =-w_1 \vc u \cdot (\vc s_1 \times \vc s_2).
\]
Since $\vc s_0 + \vc s_1 + \vc s_2=0$ we have $ \vc s_1 \times \vc s_2 = \vc s_2 \times \vc s_0 = \vc s_0 \times \vc s_1$ and thus
\[
   \pi_p \odot \pi_{s_i} = -w_i \vc u \cdot (\vc v_1 \times \vc v_2).
\]
Having the barycentric coordinates of the intersection on the triangle we can compute also its local coordinate on $p$ from its parametric form:
\[
   X_i  = A_i + t u_i, \text{ for } i=1,2,3
\]
We use $i$ with maximal $|u_i|$ for practical computation.

Next, we shall discuss the case when some of the permuted inner products are zero. In fact, we test if any barycentric coordinate is under given tolerance 
($\abs{w_i} \le \epsilon$). Various \emph{pathological} cases are detected and further topological 
information is atached to the intersection data.  In the case of one zero coordinate (non-coplanar intersection on the edge)
or two zero coordinates (intersection in the vertex), we add the local edge index or the local vertex index to the intersection, respectively.
When all coordinates are (close to) zero, the line and the triangle are (nearly) coplanar. For every edge $s_i$ we compute
its intersection with the line segment $p$ in terms of the local coordinates $t_i$ and $t_p$. Only intersections with $t_i \in (-\epsilon, 1+\epsilon)$
are accepted, for $t_i$ close to $0$ and $1$ we append the vertex index to the intersection. Edges parallel to $p$ or lying on the $p$ are skipped.
Finally, we obtain zero, one, or two intersection points with additional topological information.

>>>>

\subsection{Intersection Line-Triangle (1D-2D)}
\notePE{In this paragraph we investigate real lines. Trimming comes later when tolerance is denoted, see next PE note.}
Let us consider a line $p$ with parametric equation $\vc X = \vc A + t\vc u$, $t\in\R$ and a triangle given by vertices $(\vc V_0, \vc V_1, \vc V_2)$ 
with oriented sides $s_i=(\vc V_j, \vc V_k)$, $j=(i+1)\text{ mod }3$, $k=(i+2)\text{ mod }3$. The permuted inner products $\pi_p \odot \pi_{s_i},\, i=0,1,2$ have the same sign if and only if there
is an intersection point with the       line $p$ inside the triangle. In such a case, the inner products provides scaled barycentric coordinates of the intersection on the triangle.
In particular for the barycentric coordinate $w_i$ related to the vertex $\vc V_i$ we have
\[
   w_i = \frac{\pi_p \odot \pi_{s_i}}{ \sum_{j=0}^{2} \pi_p \odot \pi_{s_i} }
\]
Indeed, using the barycentric coordinates the intersection point can be expressed as $\vc X = \vc V_0 + w_1 \vc s_2 - w_2 \vc s_1$.
The line $p$ have \plucker coordinates $(\vc u, \vc u \times \vc X)$ (\plucker coordinates are invariant to change of initial point). Combining these two expressions we get
\[
   \pi_p \odot \pi_{s_1} = \vc u \cdot (\vc s_1 \times \vc V_2) + \vc s_1 \cdot ( \vc u \times [\vc V_0 + w_1\vc s_2 - w_2 \vc s_1])
   =-w_1 \vc u \cdot (\vc s_1 \times \vc s_2).
\]
Since $\vc s_0 + \vc s_1 + \vc s_2=0$ we have $ \vc s_1 \times \vc s_2 = \vc s_2 \times \vc s_0 = \vc s_0 \times \vc s_1$ and thus
\[
   \pi_p \odot \pi_{s_i} = -w_i \vc u \cdot (\vc v_1 \times \vc v_2).
\]
Having the barycentric coordinates of the intersection on the triangle we can compute also its local coordinate on $p$ from its parametric form:
\[
   X_i  = A_i + t u_i, \text{ for } i=1,2,3
\]
We use $i$ with maximal $|u_i|$ for practical computation.

Next, we shall discuss the case when some of the permuted inner products are zero. In fact, we test if any barycentric coordinate is under given tolerance 
($\abs{w_i} \le \epsilon$). Various \emph{pathological} cases are detected and further topological 
information is atached to the intersection data.  In the case of one zero coordinate (non-coplanar intersection on the edge)
or two zero coordinates (intersection in the vertex), we add the local edge index or the local vertex index to the intersection, respectively.
When all coordinates are (close to) zero, the line and the triangle are (nearly) coplanar. For every edge $s_i$ we compute
its intersection with the line segment $p$ in terms of the local coordinates $t_i$ and $t_p$. Only intersections with $t_i \in (-\epsilon, 1+\epsilon)$
are accepted, for $t_i$ close to $0$ and $1$ we append the vertex index to the intersection. Edges parallel to $p$ or lying on the $p$ are skipped.

Finally, we obtain zero, one, or two intersection points with additional topological information.
The intersection point data structure keeps for each of the two intersecting elements:
\begin{itemize}
 \item Barycentric coordinate of the intersection.
 \item Dimension of the most specific subelement (face, edge, vertex) where the point lies.
 \item Local index of the most specific subelement.
\end{itemize}
Further, we store the sign of the permuted inner product in order to detect various special and degenerate cases higher dimensional cases.
